Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints\thanks{This research was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2020-03-29 2019/NUMTA_Barkalov/SoftComputing/article.tex
fix some reviewer comments

	\textbf{Statement}. Let us assume that the current minimum value $z_M^\ast$  from (\ref{z_estimates}) at the $k^{th}$ iteration is achieved at the boundary point of some interval $[x_{i-1}, x_i], i=i(k)$. Then, $R_{glob}(i) \geq R_{loc}(i)$.

@@

The statement is proved.

@@

	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
\begin{theorem}\label{theorem:2} Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem \ref{theorem:1} are fulfilled. Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1,\; 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e. satisfies the condition 
\begin{align}\label{phi_bar_y}
	\varphi(\bar y) & = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \nonumber \\
	& \leq \varphi(y_{\epsilon}).
\end{align}
\end{theorem}

@@

	At every iteration of the IA-DL method there will exist such interval $[x_{j-1}, x_j]$, $j=j(k)$, that $\bar x \in [x_{j-1}, x_j]$. According to our assumption that $\bar y = y(\bar x)$ is not the limit point, there will exist such number $k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$. That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ at every iteration such interval $[x_{t-1}, x_t], t=t(k)$ will be selected that $t \neq j$.

@@

In \cite{Strongin2000} it has been proven that for any limit point $x'$ of the sequence $\{x^k\}$, generated by IA while solving the problem at any $r_\nu > 1, \; 1 \leq \nu \leq m+1$, there exists a sequence
of the intervals $[x_{t-1}, x_t], \; t=t(k)$, contracting to the point $x'$, i.e. $x' \in [x_{t-1}, x_t], \; t=t(k)$. Characteristics $R(t)$ of the intervals of the specified sequence will be positive and $R(t) \rightarrow 0$ at $k \rightarrow \infty$. 
}

@@

Indeed, both $j$ and $t$ depend from $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.
However, number $j(k)$ corresponds to the same interval with fixed characteristic \linebreak $R_{glob}(j(k))>0$, whereas number $t(k)$ corresponds to different intervals with different characteristics \linebreak $R_{loc}(t(k))>0$, where $R_{loc}(t(k)) \rightarrow 0$. From this we can conclude that the condition $R_{glob}(j) > \rho R_{loc}(t)$ is fulfilled for a large enough value of $k$.
}

>>>>

	\textbf{Lemma 1}. Let us assume that the current minimum value $z_M^\ast$  from (\ref{z_estimates}) at the $k^{th}$ iteration is achieved at the boundary point of some interval $[x_{i-1}, x_i], i=i(k)$. Then, $R_{glob}(i) \geq R_{loc}(i)$.

@@

The lemma is proved.

@@

	Let us formulate and prove the convergence conditions for the new algorithm IA-DL.\\
	{ \color{red}

	\textbf{Lemma 2}. 
Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and the conditions 1--3 of Theorem \ref{theorem:1} are fulfilled.
Let $\bar{y}=y(\bar{x})$ be a limit point of the trial sequence $\{y^k\} = \{y(x^k)\}$ generated by the IA for the problem (\ref{problem}) with the parameters $r_{\nu}>1, 1 \leq \nu \leq m+1$.
Then there will be such a sequence of intervals (corresponding to iterations with numbers $q^i, \; i=1,2,...$)
\begin{equation}\label{intervals_t}
\{[x_{t-1},x_t]\}, \; t=t(q^i), \; i=1,2,...,
\end{equation}
satisfying the conditions
\begin{align}
&\bar{x} \in \bigcap_{i=1}^\infty [x_{t-1},x_t],\label{nested_int} \\ 
&\lim_{i\rightarrow\infty}\Delta_t = 0,\label{lim_delta_t} \\ 
&R(t(q^i))>0, \; i=1,2,...,\label{R_g_zero} \\ 
&\lim_{i\rightarrow\infty} R(t(q^i)) = 0, \label{lim_R}  
\end{align}
where $\Delta_t=(x_t-x_{t-1})^{1/N}$,  $R(t)$ and $t$ are from (\ref{R_1})--(\ref{R_3}) and (\ref{MaxR}) respectively.

\textbf{Proof}. As long as $\bar{x}$ is a limit point, there should exist a sequence of trial numbers $q^1, q^2,...$ meeting the conditions
\[
\bar{x} \in [x_{t-1}, x_t], \; x^{k+1} \in [x_{t-1}, x_t], \; t=t(k), k=q^1, q^2, ...,
\]
meaning that the interval $[x_{t-1}, x_t]$ containing the limit point $\bar{x}$ at the step $k$ is hit by the trial at the subsequent step $k+1$.

From Step 3 and Step 7 of the index algorithm follows that
\begin{equation}\label{max_gamma}
\max \{x_t-x^{k+1},x^{k+1}-x_{t-1}\} \leq \gamma (x_t-x_{t-1}),
\end{equation}
where $\gamma = \frac{r+1}{2r}<1$  and $ r=\min\{r_\nu:1\leq\nu\leq m+1\}>1$.
The contraction property (\ref{lim_delta_t}) and the property (\ref{nested_int}) for the product of all intervals (\ref{intervals_t}) are the consequences of (\ref{max_gamma}).

As shown in Lemma 1, at each iteration there exists an interval with strictly positive characteristic. Due to the rule (\ref{MaxR}) this leads to the conclusion that
intervals from (\ref{intervals_t}) should also have positive characteristics because they are hit by subsequent trials; i.e., the conditions (\ref{R_g_zero}) are necessary.

For intervals with indexes $\nu(x_{i-1}) = \nu(x_{i}) = \nu$ from the second condition of Theorem \ref{theorem:1} and (\ref{K_leq_L}) follows that
\[
\frac{\left|z_i-z_{i-1}\right|}{\Delta_i}\leq 2L_\nu\sqrt{N+3},
\]
and in accordance with (\ref{z_estimates}) the inequalities 
\[
z_i = g_\nu(y(x_i))\geq z_\nu^*, \; \nu=\nu(x_i)
\]
holds. Then from the rules (\ref{R_1})--(\ref{R_3}) it follows that $R(i) \leq 2\Delta_i$, if $\nu(x_{i-1}) \neq \nu(x_{i})$, and
\[
R(i) \leq \Delta_i\left(1+\left(\frac{2L_\nu\sqrt{N+3}}{r_\nu\mu_\nu}\right)^2\right), 
\]
if $\nu(x_{i-1}) = \nu(x_{i}) = \nu$. These inequalities combined with (\ref{lim_delta_t}) lead us to the validity of (\ref{lim_R}).

}

\begin{theorem}\label{theorem:2} Let us assume the index algorithm (IA) is applied to solve the problem (\ref{problem}) and for the values $r_{\nu}=r_{\nu}^{glob}, 1 \leq \nu \leq m+1,$ the conditions of Theorem \ref{theorem:1} are fulfilled. Then, the algorithm with dual estimate of the Lipschitz constants (IA-DL) with the parameters $r_{\nu}^{glob}>r_{\nu}^{loc}>1,\; 1\leq \nu \leq m+1,$ when solving the same problem also generates a sequence of the trial points $\{y^k\}=\{y(x^k)\}$, any limit point $\bar{y}$ of which belongs to the set $Y_{\epsilon}$ from (\ref{Q_intervals}), i.e. satisfies the condition 
\begin{align}\label{phi_bar_y}
	\varphi(\bar y) & = \inf\left\{\varphi(y^k):g_j(y^k) \leq 0, 1 \leq j \leq m, k = 1,2,\ldots \right\} \nonumber \\
	& \leq \varphi(y_{\epsilon}).
\end{align}
\end{theorem}

@@

	At every iteration of the IA-DL method there will exist such interval $[x_{j-1}, x_j]$, $j=j(k)$, that $\bar x \in [x_{j-1}, x_j]$. According to our assumption that $\bar y = y(\bar x)$ is not the limit point, there will exist such number $\bar k$, starting from which the trial points will not fall into the interval $[x_{j-1}, x_j]$. That is, to conduct a new trial at the point $x^{k+1} \in [x_{t-1}, x_t]$ for each $k\geq \bar k$ such interval $[x_{t-1}, x_t], t=t(k)$ will be selected that $t(k) \neq j(k)$.

@@

In Lemma 2 it has been proven that for any limit point $x'$ of the sequence $\{x^k\}$, generated by IA while solving the problem at any $r_\nu > 1, \; 1 \leq \nu \leq m+1$, there exists a sequence
of the intervals $[x_{t-1}, x_t], \; t=t(k)$, contracting to the point $x'$. Characteristics $R(t)$ of the intervals of the specified sequence will be positive and $R(t) \rightarrow 0$ at $k \rightarrow \infty$. 
}

@@

Indeed, both $j$ and $t$ depend from $k$, as the numbering of the trial points (by the lower index) changes at every iteration and the numbering of the intervals changes respectively.
%new text begin
However, $R_{glob}(j(k))>0$ is fixed for each $k \geq \bar k$ because $j(k)$ corresponds to the same interval for $k \geq \bar k$, whereas number $t(k)$ corresponds to different intervals with different characteristics $R_{loc}(t(k)) \rightarrow 0$. From this we can conclude that the condition $R_{glob}(j) > \rho R_{loc}(t)$ is fulfilled for a large enough value of $k$.
%new text end
}
